---
title: "FF-maping-project_Statistical-analysis"
author: "Valentine Thomas"
date: "2023-02-20"
output: html_document
---

// Add notes about scripts here for ease of use
!find how to set working directory for the whole document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# package to load
! Conflicts between the dplyr and plyr packages (dplyr need to be loaded before plyr as of now)
!coin conflicts solved by detaching, PairedData is still a problem

```{r}
library(ggplot2) 
library(ggpubr)
library(rstatix) 
library(dplyr)
library(plyr)
library(tidyr)
library(matrixStats)

```

#Functions for analysis

Med.sub => between [-60, -39] an [-18,-6]

```{r}
Med.subset=function(dataframe){
matrix=data.matrix(dataframe)
med.early=rowMedians(matrix,cols=5:12, na.rm=TRUE)
id1=which(med.early=='NaN')
med.early[id1]<-'NA'
med.early=as.numeric(med.early)
med.early=as.data.frame(med.early)

med.late=rowMedians(matrix, cols=19:23, na.rm=TRUE)
id3=which(med.late=='NaN')
med.late[id3]<-'NA'
med.late=as.numeric(med.late)
med.late=as.data.frame(med.late)
df=cbind(dataframe,med.early,med.late)
return(df)
}

index_rep=function(df){
  id=list()
  for (i in 1:n_distinct(df$Replicat)) { 
    id[[i]]<-assign(paste0("index", i), which(df$Replicat==i))
  }
  return(id)
}

box.rep=function(df,title){
  data<-df %>% select(Event, Replicat,med.early,med.late) %>%
  pivot_longer(., cols = c(med.early,med.late), names_to = "Var", values_to = "Val") 
  data$Replicat<-as.factor(data$Replicat)
  ggboxplot(data,"Var", "Val", color = "Replicat", palette=c("#87CEFA", "#FFDEAD"), title= title, ylab="Median distances",xlab="fusion moment", add="jitter")
}

test.norm=function(df){
print(ddply(df, .(Replicat), summarise, ntest =
   shapiro.test(med.early)))
print(ddply(df, .(Replicat), summarise, ntest =
   shapiro.test(med.late)))}

wil.test=function(df,id_list){
  data<-df %>% select(Event, Replicat,med.early, med.late) %>%
  pivot_longer(., cols = c(med.early,med.late), names_to = "Var", values_to = "Val") 
  data$Replicat<-as.factor(data$Replicat)
  ct=1
  for (i in seq_along (id_list)){
  dataR<-filter(data, Replicat == ct)
  results<-wilcox_test(dataR,Val ~ Var,paired = TRUE)
  print(results)
  ct=ct+1
    }
}

wil.test.tot=function(df){
  data<-df %>% select(Event,med.early, med.late) %>%
  pivot_longer(., cols = c(med.early,med.late), names_to = "Var", values_to = "Val") 
  results<-wilcox_test(data,Val ~ Var,paired = TRUE)
  print(results)

    
}
#needs pairedData library => conflict
paired_plots=function(df, id_list){
  n=length(id_list)
  par(mfrow=c(1,3))
  for (i in seq_along(id_list)){
  med.early<-subset(df,  Replicat == i, med.early,drop = FALSE)
 
  med.late<-subset(df,  Replicat == i, med.late, drop = FALSE)
  pd3<- paired(med.early, med.late)
  pp3<-plot(pd3, type = "profile")
  print(pp3)
  } 
}



```




#Fus1
## prepare dataset

```{r}
Fus1<-read.csv('~/Analysis/POI_replicats/Fus1/aligned_trajectories/Data/dir_distance_P_POI.csv')
col=c("Event","Protein", "Replicat", "Dist",-60,-57, -54, -51, -48, -45,-42, -39, -36,-33,-30, -27, -24, -21, -18, -15, -12,-9, -6, -3, 0)
colnames(Fus1)<-col
Fus1=Med.subset(Fus1)
Rep.id.Fus1<-index_rep(Fus1)
```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Fus1)
independence_test( med.late~ Replicat, data = Fus1)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Fus1,"Fus1")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}

#Shapiro_Wilk test: test normality assumption
test.norm(Fus1) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Fus1,Rep.id.Fus1)

wil.test.tot(Fus1)
```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Fus1,Rep.id.Fus1)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```


#Exo70
## prepare dataset

```{r}
Exo70<-read.csv('~/Analysis/POI_replicats/Exo70/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Exo70)<-col
Exo70=Med.subset(Exo70)
Rep.id.Exo70<-index_rep(Exo70)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Exo70)
independence_test( med.late~ Replicat, data = Exo70)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
library(dplyr)
box.rep(Exo70,"Exo70")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}

#Shapiro_Wilk test: test normality assumption
test.norm(Exo70) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Exo70,Rep.id.Exo70)

wil.test.tot(Exo70)
```

## Paired representation
```{r, echo=FALSE}
library(PairedData)
paired_plots(Exo70,Rep.id.Exo70)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```


#Exo84
## prepare dataset

```{r}

Exo84<-read.csv('~/Analysis/POI_replicats/Exo84/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Exo84)<-col
Exo84=Med.subset(Exo84)
Rep.id.Exo84<-index_rep(Exo84)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Exo84)

independence_test( med.late~ Replicat, data = Exo84)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Exo84, "Exo84")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}

#Shapiro_Wilk test: test normality assumption
test.norm(Exo84) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Exo84,Rep.id.Exo84)
wil.test.tot(Exo84)
```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Exo84,Rep.id.Exo84)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```


#Ypt3
## prepare dataset

```{r}

Ypt3<-read.csv('~/Analysis/POI_replicats/Ypt3/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Ypt3)<-col
Ypt3=Med.subset(Ypt3)
Rep.id.Ypt3<-index_rep(Ypt3)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Ypt3)

independence_test( med.late~ Replicat, data = Ypt3)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Ypt3, "Ypt3")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}

#Shapiro_Wilk test: test normality assumption
test.norm(Ypt3) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Ypt3,Rep.id.Ypt3)
wil.test.tot(Ypt3)
```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Ypt3,Rep.id.Ypt3)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```


#Prm1
## prepare dataset

```{r}
library(plyr)
Prm1<-read.csv('~/Analysis/POI_replicats/Prm1/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Prm1)<-col
Prm1=Med.subset(Prm1)
Rep.id.Prm1<-index_rep(Prm1)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Prm1)

independence_test( med.late~ Replicat, data = Prm1)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
library(plyr)
box.rep(Prm1, "Prm1")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}

#Shapiro_Wilk test: test normality assumption
test.norm(Prm1) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Prm1,Rep.id.Prm1)

```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Prm1,Rep.id.Prm1)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

#Dni1
## prepare dataset

```{r}
Dni1<-read.csv('~/Analysis/POI_replicats/Dni1/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Dni1)<-col
Dni1=Med.subset(Dni1)
Rep.id.Dni1<-index_rep(Dni1)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Dni1)

independence_test( med.late~ Replicat, data = Dni1)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Dni1,"Dni1")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}

#Shapiro_Wilk test: test normality assumption
test.norm(Dni1) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Dni1,Rep.id.Dni1)

```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Dni1,Rep.id.Dni1)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

#Dni2
## prepare dataset

```{r}
Dni2<-read.csv('~/Analysis/POI_replicats/Dni2/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Dni2)<-col
Dni2=Med.subset(Dni2)
Rep.id.Dni2<-index_rep(Dni2)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Dni2)

independence_test( med.late~ Replicat, data = Dni2)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Dni2,"Dni2")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}
#Shapiro_Wilk test: test normality assumption
test.norm(Dni2) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Dni2,Rep.id.Dni2)

```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Dni2,Rep.id.Dni2)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

#Cdc8
## prepare dataset

```{r}
Cdc8<-read.csv('~/Analysis/POI_replicats/Cdc8/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Cdc8)<-col
Cdc8=Med.subset(Cdc8)
Rep.id.Cdc8<-index_rep(Cdc8)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Cdc8)

independence_test( med.late~ Replicat, data = Cdc8)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Cdc8, "Cdc8")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}
#Shapiro_Wilk test: test normality assumption
test.norm(Cdc8) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Cdc8,Rep.id.Cdc8)

```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Cdc8,Rep.id.Cdc8)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```

#Myo51
## prepare dataset

```{r}
Myo51<-read.csv('~/Analysis/POI_replicats/Myo51/aligned_trajectories/Data/dir_distance_P_POI.csv')
colnames(Myo51)<-col
Myo51=Med.subset(Myo51)
Rep.id.Myo51<-index_rep(Myo51)

```

## test replicats
```{r}
# Permutation test: test homogeneite replicats
library(coin)
independence_test( med.early~ Replicat, data = Myo51)

independence_test( med.late~ Replicat, data = Myo51)
detach("package:coin", unload=TRUE)
```

##boxplot representation
```{r , echo=FALSE}
box.rep(Myo51, "Myo51")
```

## Statistical tests

```{r echo=TRUE, paged.print=FALSE, results='markup'}
#Shapiro_Wilk test: test normality assumption
test.norm(Myo51) #data do not follow normal distribution

# Wilcoxon signed rank test: use of non parametric test !assumption distances between group are symmetric= might not be true
wil.test(Myo51,Rep.id.Myo51)

```

## Paired representation
```{r , echo=FALSE}
library(PairedData)
paired_plots(Myo51,Rep.id.Myo51)
detach("package:PairedData", unload=TRUE)
detach("package:MASS", unload=TRUE)
```


