---
title: "FF mapping project_Statistics"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache.rebuild = TRUE)
```
*Ctrl+Alt+I*
#useful packages
```{r}
library(ggplot2) 
library(ggpubr)
library(rstatix) 
library(dplyr)
library(plyr)
library(tidyr)
library(matrixStats)
```


charge the datasets 

```{r}
setwd('~/Analysis/POI_replicats/')

#medium group
Fus1<-read.csv('Fus1/aligned_trajectories/Data/dir_distance_P_POI.csv')
Exo70<-read.csv('Exo70/aligned_trajectories/Data/dir_distance_P_POI.csv')
Exo84<-read.csv('Exo84/aligned_trajectories/Data/dir_distance_P_POI.csv')

#Proximal group
Prm1<-read.csv('Prm1/aligned_trajectories/Data/dir_distance_P_POI.csv')
Dni1<-read.csv('Dni1/aligned_trajectories/Data/dir_distance_P_POI.csv')
Dni2<-read.csv('Dni2/aligned_trajectories/Data/dir_distance_P_POI.csv')

#Distal group
Cdc8<-read.csv('Cdc8/aligned_trajectories/Data/dir_distance_P_POI.csv')
Myo51<-read.csv('Myo51/aligned_trajectories/Data/dir_distance_P_POI.csv')

col=c("Event","Protein", "Replicat", "Dist",-60,-57, -54, -51, -48, -45,-42, -39, -36,-33,-30, -27, -24, -21, -18, -15, -12,-9, -6, -3, 0)
colnames(Fus1)<-col
colnames(Exo70)<-col
colnames(Exo84)<-col

colnames(Prm1)<-col
colnames(Dni1)<-col
colnames(Dni2)<-col

colnames(Cdc8)<-col
colnames(Myo51)<-col


```

#Prepare datasets for statistical test
need 3 subgroups containing medians [-60,-30[ (-60 to -30), [-40,-20] (-27 to -9, [-20,0](-6 to -0) for each event of each replicate 


```{r}
Med.subset=function(dataframe)
{
matrix=data.matrix(dataframe)
med.early=rowMedians(matrix,cols=5:15, na.rm=TRUE)
id1=which(med.early=='NaN')
med.early[id1]<-'NA'
med.early=as.numeric(med.early)
med.early=as.data.frame(med.early)
med.mid=rowMedians(matrix,cols=16:22, na.rm=TRUE)
id2=which(med.mid=='NaN')
med.mid[id2]<-'NA'
med.mid=as.numeric(med.mid)
med.mid=as.data.frame(med.mid)
med.late=rowMedians(matrix, cols=23:25, na.rm=TRUE)
id3=which(med.late=='NaN')
med.late[id3]<-'NA'
med.late=as.numeric(med.late)
med.late=as.data.frame(med.late)
df=cbind(dataframe,med.early,med.mid,med.late)

return(df)
}

#Mid grouo
Fus1=Med.subset(Fus1)
Exo70=Med.subset(Exo70)
Exo84=Med.subset(Exo84)

#Proximal group
Prm1=Med.subset(Prm1)
Dni1=Med.subset(Dni1)
Dni2=Med.subset(Dni2)

#Distal group
Cdc8=Med.subset(Cdc8)
Myo51=Med.subset(Myo51)
```
# Statistical tests

getting id of event from the different replicates

```{r}
# compare distribution for each POI between each replicates
index_rep=function(df){
  id=list()
  for (i in 1:n_distinct(df$Replicat)) { 
    id[[i]]<-assign(paste0("index", i), which(df$Replicat==i))
  }
  return(id)
}

#mid group
Rep.id.Fus1<-index_rep(Fus1)
Rep.id.Exo70<-index_rep(Exo70)
Rep.id.Exo84<-index_rep(Exo84)

#Proximal group
Rep.id.Prm1<-index_rep(Prm1)
Rep.id.Dni1<-index_rep(Dni1)
Rep.id.Dni2<-index_rep(Dni2)

#Distal group
Rep.id.Myo51<-index_rep(Myo51)
Rep.id.Cdc8<-index_rep(Cdc8)
```

# Graphical representation
boxplot of medians at different fusion stage, all and by replicats
Notes: could be interesting to look at violin plots too => asks Melvin if he knows how to do them 
```{r}

box.rep=function(df){
  data<-df %>% select(Event, Replicat,med.early,med.mid,med.late) %>%
  pivot_longer(., cols = c(med.early,med.mid,med.late), names_to = "Var", values_to = "Val") 
  data$Replicat<-as.factor(data$Replicat)
  ggboxplot(data,"Var", "Val", color = "Replicat", palette=c("#87CEFA", "#FFDEAD"))
}

box.rep(Fus1)
box.rep(Exo70)
box.rep(Exo84)

box.rep(Prm1)
box.rep(Dni1)
box.rep(Dni2)

box.rep(Cdc8)
box.rep(Myo51)
```

#Paired t-test => each median is dependant on previous one
most of the data do not follow normal distribution => go to non parametric tests
Wilcoxon signed ranks test


```{r echo=TRUE, paged.print=FALSE, results='markup'}
#test data normality
test.norm=function(df){
print(ddply(df, .(Replicat), summarise, ntest =
   shapiro.test(med.early)))
print(ddply(df, .(Replicat), summarise, ntest =
   shapiro.test(med.mid)))
print(ddply(df, .(Replicat), summarise, ntest =
   shapiro.test(med.late)))}

test.norm(Fus1)
test.norm(Exo70)
test.norm(Exo84)

test.norm(Prm1)
test.norm(Dni1)
test.norm(Dni2)

test.norm(Cdc8)
test.norm(Myo51)


 
```

#Wilcoxon signed rank test

Note: might not be the right test either => assume symmetry diferrence between group: might not be the case here 

```{r}
wil.test=function(df,id_list){
  data<-df %>% select(Event, Replicat,med.early,med.mid, med.late) %>%
  pivot_longer(., cols = c(med.early,med.mid,med.late), names_to = "Var", values_to = "Val") 
  data$Replicat<-as.factor(data$Replicat)
  ct=1
  for (i in seq_along (id_list)){
  dataR<-filter(data, Replicat == ct)
  results<-wilcox_test(dataR,Val ~ Var,paired = TRUE)
  print(results)
  ct=ct+1
    }
}

wil.test(Fus1,Rep.id.Fus1)
wil.test(Exo70,Rep.id.Exo70)
wil.test(Exo84,Rep.id.Exo84)

wil.test(Prm1,Rep.id.Prm1)
wil.test(Dni1,Rep.id.Dni1)
wil.test(Dni2,Rep.id.Dni2)

wil.test(Cdc8,Rep.id.Cdc8)
wil.test(Myo51,Rep.id.Myo51)

```
## Graphic representation 
Plot data as paired graphs

```{r echo=TRUE, paged.print=TRUE, results='markup'}
library(PairedData) # /!\ library seems to interfer with dyplr select()
paired_plots=function(df, id_list){
  n=length(id_list)
  par(mfrow=c(1,3))
  for (i in seq_along(id_list)){
  med.early<-subset(df,  Replicat == i, med.early,drop = FALSE)
  med.mid<-subset(df,  Replicat == i, med.mid, drop = FALSE)
  med.late<-subset(df,  Replicat == i, med.mid, drop = FALSE)
  pd <- paired(med.early, med.mid)
  pd2<- paired(med.mid,med.late)
  pd3<- paired(med.early, med.late)
  pp1<-plot(pd, type = "profile")
  pp2<-plot(pd2, type = "profile")
  pp3<-plot(pd3, type = "profile")
  print(pp1)
  print(pp2)
  print(pp3)
  } 
}

paired_plots(Fus1,Rep.id.Fus1)
paired_plots(Exo70,Rep.id.Exo70)
paired_plots(Exo84,Rep.id.Exo84)

paired_plots(Prm1,Rep.id.Prm1)
paired_plots(Dni1,Rep.id.Dni1)
paired_plots(Dni2,Rep.id.Dni2)

paired_plots(Cdc8,Rep.id.Cdc8)
paired_plots(Myo51,Rep.id.Myo51)

```

# Permutation tests
diff premutation tests: of independence (is there a difference between value from indvd from group 1 and group2) and of symmetry (is there a difference between the value 1 and 2 of an individual)=> could replace wilcoxon test? 



```{r}
# permutation test : independance test
library(coin)

#test med of early tp

independence_test( med.early~ Replicat, data = Fus1)
independence_test( med.early~ Replicat, data = Exo70)
independence_test( med.early~ Replicat, data = Exo84)

independence_test( med.early~ Replicat, data = Prm1)
independence_test( med.early~ Replicat, data = Dni1)
independence_test( med.early~ Replicat, data = Dni2)

independence_test( med.early~ Replicat, data = Cdc8)
independence_test( med.early~ Replicat, data = Myo51)

```
```{r}
#test med of medium time points
independence_test( med.mid~ Replicat, data = Fus1)
independence_test( med.mid~ Replicat, data = Exo70)
independence_test( med.mid~ Replicat, data = Exo84)

independence_test( med.mid~ Replicat, data = Prm1)
independence_test( med.mid~ Replicat, data = Dni1)
independence_test( med.mid~ Replicat, data = Dni2)

independence_test( med.mid~ Replicat, data = Cdc8)
independence_test( med.mid~ Replicat, data = Myo51)

```

```{r}
#test med of late time points
independence_test( med.late~ Replicat, data = Fus1)
independence_test( med.late~ Replicat, data = Exo70)
independence_test( med.late~ Replicat, data = Exo84)

independence_test( med.late~ Replicat, data = Prm1)
independence_test( med.late~ Replicat, data = Dni1)
independence_test( med.late~ Replicat, data = Dni2)

independence_test( med.late~ Replicat, data = Cdc8)
independence_test( med.late~ Replicat, data = Myo51)

```

Issues: some replicats are significantly different 


# Is fus1 significantly different from Myo52 in later timepoints?
=> one sample wilc


